---
title: "DS Encoding & Recognition — Behavioral + ET"
format: html
editor: source
---

## Overview

End-to-end pipeline for the DS memory-context eye-tracking study. Two
phases:

- **Encoding** — participants view background scenes paired with an object
  on either the left or right side. Each composite stim filename encodes
  `<background>.<object>.<emo>.<location>.jpeg`.
- **Recognition** — participants make old/new judgments on backgrounds.
  Eye tracking is recorded only during the background-task.

Two participant groups: **TD** (typically developing, in `Data/TD/`) and
**DS** (Down syndrome, mirrored layout in `Data/DS/`). This doc currently
runs against TD; flip the paths in the `paths` chunk to run DS.

**Conventions used throughout**

- Screen: 1920 × 1080 px, y grows downward in screen coords.
- Picture displayed at 700 × 550 px centered on screen → bounds
  x: 610–1310, y: 265–815. Left/Right AOIs split that box at x = 960.
- Tobii `system_time_stamp` is microseconds → divide by 1000 for ms.
- I-VT params match Tobii Pro Lab defaults (30°/s, 60 ms min fixation,
  75 ms merge gap, 0.5° merge angle); `one_degree = 40 px/deg` assumes
  ~65 cm viewing on a 24″ 1920×1080 monitor — recompute for your setup.

**Key outputs**

| object | what it is |
|---|---|
| `encoding`, `recognition` | per-trial behavioral tables |
| `gaze_trial`, `recognition_gaze_trial` | gaze samples assigned to trials |
| `fixations`, `recognition_fixations` | I-VT fixation events |
| `fix_labeled` | encoding fixations + AOI label (Left / Right / Outside) |
| `recognition_acc`, `recognition_acc_by_condition` | hit/FA/d′/c |
| `validation`, `recognition_validation` | per-validation calibration metrics |
| `reinstatement` | encoding↔recognition fixation-pattern similarity |

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
```

## Paths

Each data type lives in its own folder. Files are matched by participant ID
parsed from the filename (`^DS\d+[_-]\d+` — IDs use either `_` or `-`).

```{r paths}
# TD = typically developing. Mirror this layout under Data/DS/ for the
# Down syndrome group when those data come in.
td_behavioral_dir <- here("Data", "TD", "Behavioral")
td_et_dir         <- here("Data", "TD", "Eye_tracking")

behavioral_dir <- file.path(td_behavioral_dir, "encoding")
gaze_dir       <- file.path(td_et_dir,         "encoding",    "raw_gaze_msg")
msg_dir        <- gaze_dir

recognition_behavioral_dir <- file.path(td_behavioral_dir, "recognition")
recognition_gaze_dir       <- file.path(td_et_dir,         "recognition", "raw_gaze_msg")
recognition_msg_dir        <- recognition_gaze_dir

# Group label gets stamped into the output directory so TD and DS results
# stay separate. Change to "DS" when running the DS group through.
group_label <- "TD"
output_dir  <- here("output", group_label)
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

screen_w_px <- 1920
screen_h_px <- 1080

extract_pid <- function(path) str_extract(basename(path), "^DS\\d+[_-]\\d+")
```

## Loaders

Each loader takes a single file path and returns a tibble tagged with
`participant`. The pipelines below map them over every file in the
corresponding folder.

| function | reads | returns |
|---|---|---|
| `read_behavioral` | encoding PsychoPy CSV | one row per trial |
| `read_recognition_behavioral` | recognition PsychoPy CSV | one row per back-task trial (≤90) |
| `read_gaze` | Tobii gaze CSV | one row per sample, x/y in px |
| `read_msg` | Tobii msg CSV | one row per event message |
| `parse_validation` | a `msg` tibble | one row per validation block, metrics as numeric columns |
| `recognition_accuracy` | recognition tibble | per-group HR / FAR / d′ / c (log-linear corrected) |
| `downsample_gaze` | gaze tibble | gaze binned to a fixed sample rate |

```{r loaders}
read_behavioral <- function(path, n_practice = 3) {
  # Encoding starts with `n_practice` practice trials that have no ET
  # data — Tobii msg onsets start at the first real trial. Drop those
  # rows so behavioral trial 1 aligns with ET trial 1.
  read_csv(path, show_col_types = FALSE) |>
    filter(!is.na(Background), !is.na(Object), !is.na(Condition)) |>
    slice(-seq_len(n_practice)) |>
    mutate(
      participant = extract_pid(path),
      trial       = row_number(),
      phase = "encoding"
      ) |>
    select(
      participant,
      trial,
      Background,
      Object,
      Condition,
      List,
      mouse_clicked_name = mouse.clicked_name,
      phase
    )
}

read_gaze <- function(path) {
  read_csv(path, show_col_types = FALSE) |>
    select(
      system_time_stamp,
      left_gaze_point_on_display_area_x,
      left_gaze_point_on_display_area_y,
      right_gaze_point_on_display_area_x,
      right_gaze_point_on_display_area_y
    ) |>
    mutate(
      participant = extract_pid(path),
      time_ms     = system_time_stamp / 1000,
      left_x_px   = left_gaze_point_on_display_area_x  * screen_w_px,
      left_y_px   = left_gaze_point_on_display_area_y  * screen_h_px,
      right_x_px  = right_gaze_point_on_display_area_x * screen_w_px,
      right_y_px  = right_gaze_point_on_display_area_y * screen_h_px,
      avg_x_px    = rowMeans(cbind(left_x_px, right_x_px), na.rm = TRUE),
      avg_y_px    = rowMeans(cbind(left_y_px, right_y_px), na.rm = TRUE)
    ) |>
    select(
      participant,
      system_time_stamp,
      time_ms,
      left_x_px, left_y_px,
      right_x_px, right_y_px,
      avg_x_px, avg_y_px
    )
}

read_msg <- function(path) {
  read_csv(path, show_col_types = FALSE) |>
    mutate(participant = extract_pid(path)) |>
    select(participant, system_time_stamp, msg)
}

# Recognition behavioral CSV. Column names assume PsychoPy headers — adjust
# if your task spreadsheet uses different ones (e.g. `key_resp.keys`,
# `key_resp.corr`).
read_recognition_behavioral <- function(path) {
  read_csv(path, show_col_types = FALSE) |>
    # Eye tracking was recorded only during the background task; back.started
    # is populated only on those trials. Hard-cap at 90 trials in case any
    # extra rows slip through.
    filter(!is.na(back.started)) |>
    slice_head(n = 90) |>
    mutate(
      participant = extract_pid(path),
      trial       = row_number(), 
      phase = "recognition"
    ) |>
    select(
      participant,
      trial,
      Background,
      Composite,
      Object,
      Condition,
      stimulus_status,
      List,
      response,
      accuracy, 
      phase
    )
}

# Recognition-test accuracy summary. Returns one row per group with
# n_old / n_new, hit rate, false-alarm rate, overall accuracy, d-prime, and
# criterion (c). Use `groupvars` to break down by participant, Condition,
# etc. — defaults to per-participant. Stuart–Maxwell-style log-linear
# correction (Hautus 1995) is applied when HR or FAR hits 0/1.
recognition_accuracy <- function(data, groupvars = "participant",
                                 status_col = "stimulus_status",
                                 accuracy_col = "accuracy",
                                 old_label = "old",
                                 new_label = "new") {
  data |>
    group_by(across(all_of(groupvars))) |>
    summarise(
      n_old   = sum(.data[[status_col]] == old_label, na.rm = TRUE),
      n_new   = sum(.data[[status_col]] == new_label, na.rm = TRUE),
      n_hit   = sum(.data[[status_col]] == old_label &
                      .data[[accuracy_col]] == 1, na.rm = TRUE),
      n_cr    = sum(.data[[status_col]] == new_label &
                      .data[[accuracy_col]] == 1, na.rm = TRUE),
      n_total = sum(!is.na(.data[[accuracy_col]])),
      n_correct = sum(.data[[accuracy_col]] == 1, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(
      n_fa     = n_new - n_cr,
      n_miss   = n_old - n_hit,
      accuracy = n_correct / n_total,
      # log-linear correction: add 0.5 to hits/FAs and 1 to old/new totals
      hit_rate = (n_hit + 0.5) / (n_old + 1),
      fa_rate  = (n_fa  + 0.5) / (n_new + 1),
      d_prime  = qnorm(hit_rate) - qnorm(fa_rate),
      c_bias   = -0.5 * (qnorm(hit_rate) + qnorm(fa_rate))
    )
}

# Bin gaze samples to a uniform rate (default 120 Hz = ~8.333 ms bins) by
# averaging x/y within each bin. Use this to give every trial a known
# sampling rate before preprocessing + I-VT.
downsample_gaze <- function(dataframe, bin.length = 1000 / 120,
                            timevar = "timestamp",
                            aggvars = c("participant", "trial", "time_bin")) {
  dataframe <- dataframe |>
    mutate(time_bin = floor(.data[[timevar]] / bin.length) * bin.length)

  if (identical(aggvars, "none")) return(dataframe)

  dataframe |>
    group_by(across(all_of(aggvars))) |>
    summarize(
      left_x_px  = mean(left_x_px,  na.rm = TRUE),
      left_y_px  = mean(left_y_px,  na.rm = TRUE),
      right_x_px = mean(right_x_px, na.rm = TRUE),
      right_y_px = mean(right_y_px, na.rm = TRUE),
      avg_x_px   = mean(avg_x_px,   na.rm = TRUE),
      avg_y_px   = mean(avg_y_px,   na.rm = TRUE),
      .groups = "drop"
    )
}

# Parse the "validation data quality" rows in a msg tibble into one row per
# validation block, with each metric as a numeric column.
parse_validation <- function(msg_df) {
  msg_df |>
    filter(str_detect(msg, "^validation data quality")) |>
    mutate(
      Dev_L  = as.numeric(str_match(msg, "Dev_L:\\s*([0-9.]+)")[, 2]),
      Dev_R  = as.numeric(str_match(msg, "Dev_R:\\s*([0-9.]+)")[, 2]),
      RMS_L  = as.numeric(str_match(msg, "RMS_L:\\s*([0-9.]+)")[, 2]),
      RMS_R  = as.numeric(str_match(msg, "RMS_R:\\s*([0-9.]+)")[, 2]),
      LOSS_L = as.numeric(str_match(msg, "LOSS_L:\\s*([0-9.]+)")[, 2]),
      LOSS_R = as.numeric(str_match(msg, "LOSS_R:\\s*([0-9.]+)")[, 2]),
      SD_L   = as.numeric(str_match(msg, "SD_L:\\s*([0-9.]+)")[, 2]),
      SD_R   = as.numeric(str_match(msg, "SD_R:\\s*([0-9.]+)")[, 2])
    ) |>
    select(participant, system_time_stamp,
           Dev_L, Dev_R, RMS_L, RMS_R, LOSS_L, LOSS_R, SD_L, SD_R)
}

# Per-participant nearest-sample as-of merge: attach each message to the
# gaze sample whose system_time_stamp is closest. Vectorized via
# findInterval; no event-only rows are introduced. Used below for both the
# encoding and recognition gaze pipelines. Adds `match_offset_us` so you
# can see how far each message slid (median ~2 ms at 30 Hz).
nearest_msg_to_gaze <- function(gaze, msg_events) {
  if (nrow(gaze) == 0) return(gaze)
  gaze <- gaze |> arrange(system_time_stamp)
  if (nrow(msg_events) == 0) {
    return(gaze |>
             mutate(msg = NA_character_, event = NA_character_,
                    stim = NA_character_, match_offset_us = NA_real_))
  }
  msg_events <- msg_events |> arrange(system_time_stamp)
  sample_ts  <- gaze$system_time_stamp
  idx_lo     <- findInterval(msg_events$system_time_stamp, sample_ts)
  idx_lo[idx_lo == 0L] <- 1L
  idx_hi     <- pmin(idx_lo + 1L, length(sample_ts))
  d_lo       <- abs(sample_ts[idx_lo] - msg_events$system_time_stamp)
  d_hi       <- abs(sample_ts[idx_hi] - msg_events$system_time_stamp)
  nearest    <- ifelse(d_hi < d_lo, idx_hi, idx_lo)

  msg_matched <- msg_events |>
    mutate(matched_ts      = sample_ts[nearest],
           match_offset_us = sample_ts[nearest] - system_time_stamp) |>
    group_by(matched_ts) |>
    slice_min(abs(match_offset_us), n = 1, with_ties = FALSE) |>
    ungroup() |>
    select(matched_ts, msg, event, stim, match_offset_us)

  gaze |>
    left_join(msg_matched, by = c("system_time_stamp" = "matched_ts"))
}
```

# Encoding

## Behavioral

Stack every encoding behavioral CSV in `behavioral_dir` into one tibble
keyed on `(participant, trial)`. Trial number is the filtered row order
within the file — non-task rows (e.g. instructions) are dropped first by
requiring `Background`, `Object`, and `Condition` to be non-NA.

```{r behavioral}
behavioral_files <- list.files(behavioral_dir, pattern = "\\.csv$",
                               full.names = TRUE)

encoding <- map(behavioral_files, read_behavioral) |>
  list_rbind()

encoding
```

## Eye-tracking data

Load gaze samples and trial-event messages, then assign each gaze sample to a
trial using the onset/offset timestamps in the msg file. Tobii
`system_time_stamp` values are in microseconds, so we convert to ms.

```{r load-et}
gaze_files <- list.files(gaze_dir, pattern = "_gaze\\.csv$", full.names = TRUE)
msg_files  <- list.files(msg_dir,  pattern = "_msg\\.csv$",  full.names = TRUE)

gaze <- map(gaze_files, read_gaze) |> list_rbind()
msg  <- map(msg_files,  read_msg)  |> list_rbind()
```

Pull the `validation data quality` rows out of the msg file and parse each
metric into its own column. Dev/RMS/SD are in degrees of visual angle,
LOSS is the proportion of lost samples (0–1). One row per validation block —
sessions with multiple recalibrations produce multiple rows per participant.

```{r validation}
validation <- parse_validation(msg)
validation
```

Keep msg in long format, dropping calibration/validation rows so only
`onset_*` / `offset_*` trial markers remain.

```{r msg-long}
msg_events <- msg |>
  filter(str_detect(msg, "^(onset|offset)_")) |>
  mutate(
    event = str_extract(msg, "^(onset|offset)"),
    stim  = str_remove(msg, "^(onset|offset)_")
  ) |>
  select(participant, system_time_stamp, msg, event, stim)

msg_events
```

Merge gaze and msg by stacking on `system_time_stamp` and sorting in time
*within participant*. Then assign `trial`: it ticks up by 1 at every `onset`
and clears on `offset`, so each gaze sample carries the trial of the window
it sits in (`NA` outside trials).

```{r merge-et}
# Per-trial stim lookup (one row per trial within participant) so we can
# broadcast the stimulus jpeg across every gaze sample in that trial.
trial_stim <- msg_events |>
  filter(event == "onset") |>
  group_by(participant) |>
  arrange(system_time_stamp, .by_group = TRUE) |>
  mutate(trial = row_number()) |>
  ungroup() |>
  select(participant, trial, stim) |>
  # stim format: "<background>.<object>.<emo>.<location>.jpeg".
  # A handful of stims use "_" as the delimiter instead (e.g.
  # "airport_a_n_r.jpeg") — normalize those onto stim_split so the
  # original `stim` keeps the on-disk filename (needed for image lookup).
  mutate(stim_split = if_else(
    str_count(stim, fixed(".")) < 4,
    str_replace_all(stim, "_", "."),
    stim
  )) |>
  separate_wider_delim(
    stim_split,
    delim = ".",
    names = c("background", "object", "emo", "location", "ext"),
    cols_remove = TRUE
  ) |>
  select(participant, trial, stim, background, object, emo, location)

# Nearest-sample as-of merge per participant: each onset/offset message
# gets attached to the gaze sample whose timestamp is closest. The output
# has one row per gaze sample (no synthetic event rows), with msg / event /
# stim sparsely set on the samples that absorbed a message.
gaze_trial <- gaze |>
  group_by(participant) |>
  group_modify(~ nearest_msg_to_gaze(
    .x, filter(msg_events, participant == .y$participant)
  )) |>
  ungroup() |>
  group_by(participant) |>
  arrange(system_time_stamp, .by_group = TRUE) |>
  mutate(
    is_onset      = !is.na(event) & event == "onset",
    is_offset     = !is.na(event) & event == "offset",
    onset_run     = cumsum(is_onset),
    # offsets counted *before* the current row, so the offset row
    # itself still belongs to the trial it closes.
    offset_before = lag(cumsum(is_offset), default = 0),
    in_trial      = onset_run > offset_before,
    trial         = if_else(in_trial, onset_run, NA_integer_)
  ) |>
  ungroup() |>
  select(-stim) |>
  left_join(trial_stim, by = c("participant", "trial")) |>
  filter(!is.na(trial)) |>
  group_by(participant, trial) |>
  mutate(time_ms = time_ms - min(time_ms)) |>
  ungroup() |>
  # Truncate every trial to exactly 5 s. PsychoPy's pic-display nominal
  # duration is 5 s; trimming here means every downstream analysis sees the
  # same window even when a tail of samples leaks past offset.
  filter(time_ms <= 5000) |>
  select(
    participant,
    trial,
    time_ms,
    msg,
    event,
    match_offset_us,
    stim,
    background, object, emo, location,
    left_x_px, left_y_px,
    right_x_px, right_y_px,
    avg_x_px, avg_y_px
  )

gaze_trial
```

Sanity check: every onset has a matching offset per participant, and no
onset/offset msg row was dropped during the merge.

```{r merge-et-check}
msg_balance <- msg_events |>
  group_by(participant) |>
  summarise(
    n_onset  = sum(event == "onset"),
    n_offset = sum(event == "offset"),
    balanced = n_onset == n_offset,
    .groups  = "drop"
  )

stopifnot(all(msg_balance$balanced))
stopifnot(
  nrow(filter(gaze_trial, !is.na(event))) == nrow(msg_events)
)

msg_balance
```

## Missing-data QC

Count NA samples inside the 5-s trial window. A sample is "missing" when
the binocular average is NA on either axis — both eyes lost, or one
eye dropped and the other never recovered. Reported per trial, then
rolled up per subject so trackers / participants with high data loss are
visible at a glance.

```{r encoding-missing}
trial_missing <- gaze_trial |>
  group_by(participant, trial) |>
  summarise(
    n_samples    = n(),
    n_missing    = sum(is.na(avg_x_px) | is.na(avg_y_px)),
    prop_missing = n_missing / n_samples,
    duration_ms  = max(time_ms, na.rm = TRUE) - min(time_ms, na.rm = TRUE),
    .groups = "drop"
  )

trial_missing

subject_missing <- trial_missing |>
  group_by(participant) |>
  summarise(
    n_trials             = n(),
    total_samples        = sum(n_samples, na.rm = TRUE),
    total_missing        = sum(n_missing, na.rm = TRUE),
    mean_prop_missing    = mean(prop_missing,   na.rm = TRUE),
    median_prop_missing  = median(prop_missing, na.rm = TRUE),
    max_prop_missing     = max(prop_missing,    na.rm = TRUE),
    n_trials_over_50pct  = sum(prop_missing > 0.50, na.rm = TRUE),
    .groups = "drop"
  )

subject_missing
```


## Preprocess gaze (kollaR)

`kollaR::preprocess_gaze()` interpolates short gaps and applies a moving-mean
filter, but it expects continuous samples within a single trial — so apply
it per (participant, trial) via `group_modify()`. It also requires a column
literally named `timestamp` (in ms). The nearest-sample merge above means
every row in `gaze_trial` is already a real sample (events are merged onto
samples, not extra rows), so we can downsample directly.

```{r preprocess}
library(kollaR)

# Downsample to a uniform 120 Hz first so I-VT velocities are computed
# against a known sample rate (Tobii samples can drop, leaving gaps).
gaze_120hz <- gaze_trial |>
  rename(timestamp = time_ms) |>
  downsample_gaze(bin.length = 1000 / 120) |>
  rename(timestamp = time_bin)

gaze_pp <- gaze_120hz |>
  group_by(participant, trial) |>
  group_modify(~ preprocess_gaze(
    .x,
    xcol = "avg_x_px",
    ycol = "avg_y_px"
  )) |>
  ungroup()

gaze_pp
```

## Detect fixations (I-VT)

`kollaR::algorithm_ivt()` returns a **list** (`fixations`, `saccades`,
`filt.gaze`, `velocity`) per call, so apply it per-trial via `nest()` +
`map()` and unnest the fixation table.

Parameters set to match Tobii Pro Lab's I-VT (Fixation) filter defaults:
30°/s velocity threshold, 0.5° merge angle, 60 ms min fixation duration,
75 ms merge gap, 20 ms velocity window. `one_degree` is pixels-per-degree —
40 is a reasonable default for a Tobii Spectrum at ~65 cm viewing distance
on a 24″ 1920×1080 monitor; recompute it from your actual setup if needed.

```{r ivt}
# Wrap algorithm_ivt() so trials with no valid samples return NULL instead
# of throwing. Tobii defaults below; recompute one_degree if your setup
# differs.
safe_ivt_encoding <- possibly(
  function(d) algorithm_ivt(
    d,
    xcol                   = "avg_x_px",
    ycol                   = "avg_y_px",
    velocity.filter.ms     = 20,    # Tobii velocity window
    velocity.threshold     = 30,    # Tobii I-VT classifier (deg/s)
    min.fixation.duration  = 60,    # Tobii "discard short fixations"
    merge.ms.threshold     = 75,    # Tobii max time between fixations
    distance.threshold     = 0.5,   # Tobii max angle between fixations (deg)
    one_degree             = 40     # px per degree — adjust to your setup
  ),
  otherwise = NULL
)

ivt_results <- gaze_pp |>
  nest(.by = c(participant, trial)) |>
  mutate(ivt = map(data, safe_ivt_encoding))

ivt_failed <- ivt_results |>
  filter(map_lgl(ivt, is.null)) |>
  select(participant, trial)

ivt_failed

fixations <- ivt_results |>
  filter(!map_lgl(ivt, is.null)) |>
  mutate(fixations = map(ivt, "fixations")) |>
  select(participant, trial, fixations) |>
  unnest(fixations)

saccades <- ivt_results |>
  filter(!map_lgl(ivt, is.null)) |>
  mutate(saccades = map(ivt, "saccades")) |>
  select(participant, trial, saccades) |>
  unnest(saccades)

fixations
```

## Merge fixations with behavioral

Attach trial-level behavioral fields and stim attributes (background,
object) to every fixation, then derive `emo` / `location` from the
behavioral `Condition` column (e.g. `"neg-right"`). The stim-filename
parse misfires on the `airport_a_n_r.jpeg` form (`"n"` / `"r"` instead of
`"neg"` / `"right"`), so `Condition` is the source of truth.

```{r merge-behavior}
trial_info <- encoding |>
  left_join(distinct(gaze_trial, participant, trial,
                     stim, background, object),
            by = c("participant", "trial")) |>
  separate_wider_delim(
    Condition, delim = "-",
    names = c("emo", "location"),
    cols_remove = FALSE
  )

fixations_behav <- fixations |>
  left_join(trial_info, by = c("participant", "trial"))

fixations_behav
```

## Label AOIs

Tag every fixation as **Left**, **Right**, or **Outside** based on which
half of the 700 × 550 picture box it falls in. The first summarise() is a
sanity check on coordinate extents (should hug 0–1920 / 0–1080).
`fix_summary` collapses to one row per (participant, trial, AOI) with
fixation counts and total dwell time.

```{r aoi-label}
fixations_behav |>
  summarise(
    min_x = min(x, na.rm = TRUE),
    max_x = max(x, na.rm = TRUE),
    min_y = min(y, na.rm = TRUE),
    max_y = max(y, na.rm = TRUE),
    n_trials = n_distinct(trial)
  )

fix_labeled <- fixations_behav |>
  mutate(
    # Picture is 700x550 centered on 1920x1080 -> top-left at (610, 265).
    x_pic = x - 610,
    y_pic = y - 265,
    AOI = case_when(
      x_pic >= 0   & x_pic <  350 & y_pic >= 0 & y_pic <= 550 ~ "Left",
      x_pic >= 350 & x_pic <= 700 & y_pic >= 0 & y_pic <= 550 ~ "Right",
      TRUE ~ "Outside"
    )
  )

fix_summary <- fix_labeled |>
  filter(AOI != "Outside") |>
  group_by(participant, trial, AOI, background, object, emo, location) |>
  summarise(
    mean_fix_duration = mean(duration, na.rm = TRUE),
    n_fixations       = n(),
    total_dwell_time  = sum(duration, na.rm = TRUE),
    .groups = "drop"
  )
```

## AOI overlay sanity check

Plot the Left / Right AOI rectangles on top of an actual stimulus image to
verify the bounds match where the L / R objects are drawn. Fixations from
one trial are overlaid and colored by their assigned AOI — points should sit
inside the same-colored rectangle.

```{r aoi-overlay}
library(jpeg)

# Picture geometry (screen coords, y grows downward).
pic_x_min <- 610;  pic_x_max <- 1310
pic_y_min <- 265;  pic_y_max <- 815
pic_x_mid <- (pic_x_min + pic_x_max) / 2  # 960 — Left/Right boundary
flip_y    <- function(y) screen_h_px - y  # screen y → ggplot y-up

# Pick a (participant, trial) whose stim image is on disk.
image_dir   <- here("image")
image_files <- list.files(image_dir)
target_aoi  <- fix_labeled |>
  filter(stim %in% image_files, AOI != "Outside") |>
  slice(1)

img       <- readJPEG(file.path(image_dir, target_aoi$stim))
trial_fix <- fix_labeled |>
  filter(participant == target_aoi$participant,
         trial       == target_aoi$trial)

ggplot() +
  annotation_raster(img,
                    xmin = pic_x_min, xmax = pic_x_max,
                    ymin = flip_y(pic_y_max), ymax = flip_y(pic_y_min)) +
  annotate("rect",
           xmin = pic_x_min, xmax = pic_x_mid,
           ymin = flip_y(pic_y_max), ymax = flip_y(pic_y_min),
           fill = NA, color = "#1f77b4", linewidth = 1.2) +
  annotate("rect",
           xmin = pic_x_mid, xmax = pic_x_max,
           ymin = flip_y(pic_y_max), ymax = flip_y(pic_y_min),
           fill = NA, color = "#d62728", linewidth = 1.2) +
  annotate("text",
           x = (pic_x_min + pic_x_mid) / 2, y = flip_y(pic_y_min) + 20,
           label = "Left",  color = "#1f77b4", fontface = "bold") +
  annotate("text",
           x = (pic_x_mid + pic_x_max) / 2, y = flip_y(pic_y_min) + 20,
           label = "Right", color = "#d62728", fontface = "bold") +
  geom_point(data = trial_fix,
             aes(x = x, y = flip_y(y), color = AOI, size = duration),
             alpha = 0.7) +
  scale_color_manual(values = c(Left = "#1f77b4", Right = "#d62728",
                                Outside = "gray50")) +
  scale_size_continuous(range = c(2, 8), guide = "none") +
  coord_fixed(xlim = c(0, screen_w_px), ylim = c(0, screen_h_px)) +
  labs(
    title    = paste0(target_aoi$participant, " — trial ", target_aoi$trial),
    subtitle = target_aoi$stim,
    x = "screen x (px)", y = "screen y, flipped (px)"
  ) +
  theme_minimal()
```

## All participants on the same stimulus

Fixations on a single stim from every participant who saw it, faceted.
The stim is whichever image is on disk in `image/` — only participants
with a trial for that stim will appear.

```{r aoi-all-participants}
shared_stim <- list.files(image_dir)[1]
img_shared  <- readJPEG(file.path(image_dir, shared_stim))

shared_fix <- fix_labeled |>
  filter(stim == shared_stim)

ggplot(shared_fix, aes(x = x, y = flip_y(y),
                       color = participant, size = duration)) +
  annotation_raster(img_shared,
                    xmin = pic_x_min, xmax = pic_x_max,
                    ymin = flip_y(pic_y_max), ymax = flip_y(pic_y_min)) +
  annotate("rect",
           xmin = pic_x_min, xmax = pic_x_mid,
           ymin = flip_y(pic_y_max), ymax = flip_y(pic_y_min),
           fill = NA, color = "white", linewidth = 1) +
  annotate("rect",
           xmin = pic_x_mid, xmax = pic_x_max,
           ymin = flip_y(pic_y_max), ymax = flip_y(pic_y_min),
           fill = NA, color = "white", linewidth = 1) +
  geom_point(alpha = 0.75) +
  scale_size_continuous(range = c(2, 8), guide = "none") +
  coord_fixed(xlim = c(0, screen_w_px), ylim = c(0, screen_h_px)) +
  labs(
    title    = "Fixations across participants",
    subtitle = shared_stim,
    color    = "Participant",
    x = "screen x (px)", y = "screen y, flipped (px)"
  ) +
  theme_minimal()
```

## Animated scanpath on stimulus image

Plot fixations from a single trial overlaid on the actual stimulus image.
The picture is 700x550 centered on a 1920x1080 screen, so its bounding box
in screen pixels is x: 610-1310, y: 265-815. `kollaR::animated_fixation_plot()`
needs the gaze data with `onset`/`offset` columns (already in ms relative to
the onset of the trial) and a `background.images` table with absolute pixel
bounds.

```{r animated-plot, eval=FALSE}
# Pick a (participant, trial) whose stim image we actually have on disk.
image_dir <- here("image")
target <- fix_labeled |>
  filter(stim %in% list.files(image_dir)) |>
  slice(1)

target_pid   <- target$participant
target_trial <- target$trial
trial_stim_name <- target$stim

background.images <- data.frame(
  path    = file.path(image_dir, trial_stim_name),
  min.x   = 610,
  min.y   = 265,
  max.x   = 1310,
  max.y   = 815,
  onset   = 0,
  offset  = 6000   # picture shown for ~5.2 s; pad a bit
)

trial_fixations <- fix_labeled |>
  filter(participant == target_pid, trial == target_trial) |>
  select(x, y, onset, offset, duration)

animated_fixation_plot(
  trial_fixations,
  xres                = 1920,
  yres                = 1080,
  plot.onset          = 0,
  plot.offset         = 6000,
  background.images   = background.images,
  save.gif            = FALSE,
  n.loops             = 0,
  resolution.scaling  = 0.5,
  gif.dpi             = 150,
  show.progress       = FALSE
)
```

## Object-side dwell by emotion

Per-trial dwell on the object side vs. background side, split by emotion.
`on_object` is TRUE when the AOI matches the cued `location`.

```{r aoi-summary}
emo_dwell <- fix_labeled |>
  filter(AOI != "Outside") |>
  mutate(on_object = AOI == str_to_title(location)) |>
  group_by(participant, emo, on_object) |>
  summarise(
    n_fix       = n(),
    total_dwell = sum(duration, na.rm = TRUE),
    mean_dur    = mean(duration, na.rm = TRUE),
    .groups     = "drop_last"
  ) |>
  mutate(prop_dwell = total_dwell / sum(total_dwell)) |>
  ungroup()

emo_dwell
```

## Encoding fixations by emo × location

Break encoding fixations into the four cells of interest — **neg-left,
neg-right, neu-left, neu-right** — at the participant level. Counts and
durations are averaged across trials within each cell.

```{r encoding-emo-loc}
encoding_fix_by_emoloc <- fix_labeled |>
  group_by(participant, trial, emo, location) |>
  summarise(
    n_fixations       = n(),
    mean_fix_duration = mean(duration, na.rm = TRUE),
    total_dwell_time  = sum(duration,  na.rm = TRUE),
    .groups = "drop"
  ) |>
  group_by(participant, emo, location) |>
  summarise(
    n_trials              = n(),
    mean_n_fixations      = mean(n_fixations,       na.rm = TRUE),
    mean_fix_duration     = mean(mean_fix_duration, na.rm = TRUE),
    mean_total_dwell_time = mean(total_dwell_time,  na.rm = TRUE),
    .groups = "drop"
  )

encoding_fix_by_emoloc
```

# Recognition

Mirrors the encoding pipeline against the `recognition_*` folders. The
behavioral table keeps `Background`, `Composite`, `Object`, `Condition`,
`stimulus_status`, `cb`, `response`, and `accuracy`.

## Behavioral

```{r recognition-behavioral}
recognition_behavioral_files <- list.files(
  recognition_behavioral_dir, pattern = "\\.csv$", full.names = TRUE
)

recognition <- map(recognition_behavioral_files, read_recognition_behavioral) |>
  list_rbind()

recognition
```

## Recognition accuracy

Per-participant hit rate, false-alarm rate, overall accuracy, d-prime, and
criterion. Pass `groupvars = c("participant", "Condition")` to break it down
by encoding condition.

```{r recognition-accuracy}
recognition_acc <- recognition_accuracy(recognition)
recognition_acc

recognition_acc_by_condition <- recognition_accuracy(
  recognition, groupvars = c("participant", "Condition")
)
recognition_acc_by_condition
```

## Picture-presentation duration check

Each background-task pic should be displayed for ~5 s. Compute
`back.stopped - back.started` per trial and flag rows where PsychoPy wrote
non-numeric junk (e.g. `[np.float64(...)]` array literals).

```{r recognition-duration}
recognition_durations <- map(recognition_behavioral_files, function(p) {
  read_csv(p, show_col_types = FALSE) |>
    filter(!is.na(back.started)) |>
    slice_head(n = 90) |>
    transmute(
      participant   = extract_pid(p),
      trial         = row_number(),
      back_started  = suppressWarnings(as.numeric(back.started)),
      back_stopped  = suppressWarnings(as.numeric(back.stopped)),
      back_duration = back_stopped - back_started,
      corrupt       = is.na(back_started) | is.na(back_stopped)
    )
}) |> list_rbind()

recognition_duration_summary <- recognition_durations |>
  group_by(participant) |>
  summarise(
    n          = n(),
    n_corrupt  = sum(corrupt),
    min_dur    = min(back_duration, na.rm = TRUE),
    median_dur = median(back_duration, na.rm = TRUE),
    max_dur    = max(back_duration, na.rm = TRUE),
    mean_dur   = mean(back_duration, na.rm = TRUE),
    n_off_5s   = sum(abs(back_duration - 5) > 0.05, na.rm = TRUE),
    .groups    = "drop"
  )

recognition_duration_summary
```

## Eye-tracking data

```{r recognition-load-et}
recognition_gaze_files <- list.files(
  recognition_gaze_dir, pattern = "_gaze\\.csv$", full.names = TRUE
)
recognition_msg_files <- list.files(
  recognition_msg_dir,  pattern = "_msg\\.csv$",  full.names = TRUE
)

recognition_gaze <- map(recognition_gaze_files, read_gaze) |> list_rbind()
recognition_msg  <- map(recognition_msg_files,  read_msg)  |> list_rbind()
```

### Calibration validation

Same parse as the encoding `validation` table but for the recognition
session. Useful for spotting sessions whose data quality drifted between
phases.

```{r recognition-validation}
recognition_validation <- parse_validation(recognition_msg)
recognition_validation
```

### Trial events

```{r recognition-msg-events}
recognition_msg_events <- recognition_msg |>
  filter(str_detect(msg, "^(onset|offset)_")) |>
  mutate(
    event = str_extract(msg, "^(onset|offset)"),
    stim  = str_remove(msg, "^(onset|offset)_")
  ) |>
  select(participant, system_time_stamp, msg, event, stim)

recognition_msg_events
```

Recognition stim names aren't decomposed into background/object/emo/location
the way encoding stim names are — recognition typically shows a composite,
so we keep `stim` as one column. Adjust if your recognition filenames follow
a parseable schema.

```{r recognition-merge-et}
recognition_trial_stim <- recognition_msg_events |>
  filter(event == "onset") |>
  group_by(participant) |>
  arrange(system_time_stamp, .by_group = TRUE) |>
  mutate(trial = row_number()) |>
  ungroup() |>
  select(participant, trial, stim)

# Nearest-sample as-of merge per participant (same approach as encoding).
recognition_gaze_trial <- recognition_gaze |>
  group_by(participant) |>
  group_modify(~ nearest_msg_to_gaze(
    .x, filter(recognition_msg_events, participant == .y$participant)
  )) |>
  ungroup() |>
  group_by(participant) |>
  arrange(system_time_stamp, .by_group = TRUE) |>
  mutate(
    is_onset      = !is.na(event) & event == "onset",
    is_offset     = !is.na(event) & event == "offset",
    onset_run     = cumsum(is_onset),
    offset_before = lag(cumsum(is_offset), default = 0),
    in_trial      = onset_run > offset_before,
    trial         = if_else(in_trial, onset_run, NA_integer_)
  ) |>
  ungroup() |>
  select(-stim) |>
  left_join(recognition_trial_stim, by = c("participant", "trial")) |>
  filter(!is.na(trial)) |>
  group_by(participant, trial) |>
  mutate(time_ms = time_ms - min(time_ms)) |>
  ungroup() |>
  # Truncate every recognition trial to exactly 5 s, same as encoding.
  filter(time_ms <= 5000) |>
  select(
    participant,
    trial,
    time_ms,
    msg,
    event,
    match_offset_us,
    stim,
    left_x_px, left_y_px,
    right_x_px, right_y_px,
    avg_x_px, avg_y_px
  )

recognition_gaze_trial
```

Sanity check: every onset has a matching offset per participant, and no
onset/offset msg row was dropped during the merge.

```{r recognition-merge-et-check}
recognition_msg_balance <- recognition_msg_events |>
  group_by(participant) |>
  summarise(
    n_onset  = sum(event == "onset"),
    n_offset = sum(event == "offset"),
    balanced = n_onset == n_offset,
    .groups  = "drop"
  )

stopifnot(all(recognition_msg_balance$balanced))
stopifnot(
  nrow(filter(recognition_gaze_trial, !is.na(event))) ==
    nrow(recognition_msg_events)
)

recognition_msg_balance
```

## Missing-data QC (recognition)

Same definition as encoding: NA samples inside the 5-s trial window, per
trial and per subject.

```{r recognition-missing}
recognition_trial_missing <- recognition_gaze_trial |>
  group_by(participant, trial) |>
  summarise(
    n_samples    = n(),
    n_missing    = sum(is.na(avg_x_px) | is.na(avg_y_px)),
    prop_missing = n_missing / n_samples,
    duration_ms  = max(time_ms, na.rm = TRUE) - min(time_ms, na.rm = TRUE),
    .groups = "drop"
  )

recognition_trial_missing

recognition_subject_missing <- recognition_trial_missing |>
  group_by(participant) |>
  summarise(
    n_trials             = n(),
    total_samples        = sum(n_samples, na.rm = TRUE),
    total_missing        = sum(n_missing, na.rm = TRUE),
    mean_prop_missing    = mean(prop_missing,   na.rm = TRUE),
    median_prop_missing  = median(prop_missing, na.rm = TRUE),
    max_prop_missing     = max(prop_missing,    na.rm = TRUE),
    n_trials_over_50pct  = sum(prop_missing > 0.50, na.rm = TRUE),
    .groups = "drop"
  )

recognition_subject_missing
```

## Preprocess + I-VT

Same recipe as the encoding pipeline: downsample to a uniform 120 Hz,
interpolate short gaps with `kollaR::preprocess_gaze()`, then run I-VT
fixation detection per (participant, trial) wrapped in `possibly()` so a
single tracker-dropout trial doesn't kill the map.

```{r recognition-preprocess}
recognition_gaze_120hz <- recognition_gaze_trial |>
  rename(timestamp = time_ms) |>
  downsample_gaze(bin.length = 1000 / 120) |>
  rename(timestamp = time_bin)

recognition_gaze_pp <- recognition_gaze_120hz |>
  group_by(participant, trial) |>
  group_modify(~ preprocess_gaze(
    .x,
    xcol = "avg_x_px",
    ycol = "avg_y_px"
  )) |>
  ungroup()

recognition_gaze_pp
```

```{r recognition-ivt}
# Wrap algorithm_ivt() so trials with no valid samples return NULL instead
# of throwing. Without this, one bad trial (e.g. tracker dropout the entire
# 5 s window) kills the whole map().
safe_ivt <- possibly(
  function(d) algorithm_ivt(
    d,
    xcol                   = "avg_x_px",
    ycol                   = "avg_y_px",
    velocity.filter.ms     = 20,
    velocity.threshold     = 30,
    min.fixation.duration  = 60,
    merge.ms.threshold     = 75,
    distance.threshold     = 0.5,
    one_degree             = 40
  ),
  otherwise = NULL
)

recognition_ivt_results <- recognition_gaze_pp |>
  nest(.by = c(participant, trial)) |>
  mutate(ivt = map(data, safe_ivt))

# Trials that failed I-VT (all-NA samples or too-short data).
recognition_ivt_failed <- recognition_ivt_results |>
  filter(map_lgl(ivt, is.null)) |>
  select(participant, trial)

recognition_ivt_failed

recognition_fixations <- recognition_ivt_results |>
  filter(!map_lgl(ivt, is.null)) |>
  mutate(fixations = map(ivt, "fixations")) |>
  select(participant, trial, fixations) |>
  unnest(fixations)

recognition_saccades <- recognition_ivt_results |>
  filter(!map_lgl(ivt, is.null)) |>
  mutate(saccades = map(ivt, "saccades")) |>
  select(participant, trial, saccades) |>
  unnest(saccades)

recognition_fixations
```

## Merge fixations with behavioral

Attach the recognition behavioral fields (Background, Composite, Condition,
stimulus_status, response, accuracy) to every fixation by joining on
`(participant, trial)`.

```{r recognition-merge-behavior}
recognition_trial_info <- recognition |>
  left_join(distinct(recognition_gaze_trial, participant, trial, stim),
            by = c("participant", "trial"))

recognition_fixations_behav <- recognition_fixations |>
  left_join(recognition_trial_info, by = c("participant", "trial"))

recognition_fixations_behav
```

## Label recognition AOIs

Same picture geometry as encoding (700 × 550 box, split at x = 960), so
each recognition fixation is tagged **Left**, **Right**, or **Outside**.
At recognition we don't know which side the *originally encoded* object
sat on, but the spatial split is still meaningful for asking where the
gaze went — and it can be cross-referenced with the encoded `location`
via Background.

```{r recognition-aoi-label}
recognition_fix_labeled <- recognition_fixations_behav |>
  mutate(
    x_pic = x - pic_x_min,
    y_pic = y - pic_y_min,
    AOI = case_when(
      x_pic >= 0   & x_pic <  350 & y_pic >= 0 & y_pic <= 550 ~ "Left",
      x_pic >= 350 & x_pic <= 700 & y_pic >= 0 & y_pic <= 550 ~ "Right",
      TRUE ~ "Outside"
    )
  )

recognition_fix_labeled |> count(AOI)
```

## Recognition fixation summary

Per-trial fixation count and mean fixation duration broken down by AOI,
restricted to **correctly recognized old items**
(`stimulus_status == "old"` & `accuracy == 1`) — same filter the eyesim
pipeline uses, so analyses are scoped to the same trials. Outside-the-box
fixations are dropped before summarising.

```{r recognition-fix-summary}
recognition_fix_summary <- recognition_fix_labeled |>
  filter(stimulus_status == "old", accuracy == 1, AOI != "Outside") |>
  group_by(participant, trial, Background, Condition, AOI) |>
  summarise(
    n_fixations       = n(),
    mean_fix_duration = mean(duration, na.rm = TRUE),
    total_dwell_time  = sum(duration,  na.rm = TRUE),
    .groups = "drop"
  )

recognition_fix_summary
```

Roll the per-trial table up to one row per (participant × Condition × AOI)
— the table to feed into participant-level stats.

```{r recognition-fix-by-condition}
recognition_fix_by_condition <- recognition_fix_summary |>
  group_by(Condition, AOI) |>
  summarise(
    n_trials              = n(),
    mean_n_fixations      = mean(n_fixations,       na.rm = TRUE),
    mean_fix_duration     = mean(mean_fix_duration, na.rm = TRUE),
    mean_total_dwell_time = mean(total_dwell_time,  na.rm = TRUE),
    .groups = "drop"
  )

recognition_fix_by_condition
```

## Encoding + recognition fixations (long format)

Stack encoding and recognition fixations into one long frame keyed on
`(participant, Background)`, with `phase` distinguishing the two. Recognition
fixations are pre-filtered to `stimulus_status == "old"` and `accuracy == 1`
so only correctly recognized old items remain — accuracy/status columns are
dropped after filtering. Restricted to Backgrounds that appear in *both*
phases per participant (so every Background here has paired data).

```{r merge-recognition-encoding}
encoding_long <- fixations_behav |>
  select(participant, Background, Condition, List, phase,
         x, y, duration, onset)

recognition_long <- recognition_fixations_behav |>
  filter(stimulus_status == "old", accuracy == 1) |>
  select(participant, Background, Condition, List, phase,
         x, y, duration, onset)

paired_keys <- inner_join(
  distinct(encoding_long,    participant, Background),
  distinct(recognition_long, participant, Background),
  by = c("participant", "Background")
)

fixations_long <- bind_rows(encoding_long, recognition_long) |>
  semi_join(paired_keys, by = c("participant", "Background")) |>
  arrange(participant, Background, phase, onset)

fixations_long
```

## Gaze reinstatement (eyesim)

Use `eyesim` to test whether the recognition-phase fixation pattern
reinstates the encoding-phase pattern for the same `Background` *within
participant*. Workflow follows the package vignette:

1. Wrap `fixations_long` in `eye_table()` keyed on participant + phase +
   Background.
2. Build a fixation-density map per group with `density_by()`.
3. Run `template_similarity()` with the encoding densities as the reference
   and the recognition densities as the source. We match on a composite
   `pair_id = participant::Background` so each recognition map is compared
   to the *same participant's* encoding map for the *same* Background, and
   permute within participant for the null distribution.

Restrict to fixations that landed on the picture itself
(x: 610–1310, y: 265–815) — fixations on the surrounding gray screen are
dropped before density estimation. Coordinates stay in absolute screen
pixels so the density maps live in the same space as the raw fixations
and the AOI rectangles.

```{r eyesim-setup}
# install if needed:
# remotes::install_github("bbuchsbaum/eyesim")
library(eyesim)

fixations_long_eye <- fixations_long |>
  mutate(pair_id = paste(Background, Condition, sep = "::"))

eyetab <- eye_table(
  x = "x", y = "y",
  duration = "duration", onset = "onset",
  groupvar = c("participant", "phase", "Background", "Condition", "pair_id"),
  data    = fixations_long_eye,
  clip_bounds     = c(pic_x_min, pic_x_max, pic_y_min, pic_y_max),
  relative_coords = TRUE
)

eyetab
```

Build a fixation-density map per (participant, phase, Background) by
smoothing the (now picture-only) fixations with a Gaussian
(`sigma = 80 px` ≈ 2°) across the picture box. These per-group maps are
what `template_similarity()` compares.

```{r eyesim-density}
eyedens <- density_by(
  eyetab,
  groups  = c("participant", "phase", "Background", "Condition", "pair_id"),
  sigma   = 80,
  xbounds = c(pic_x_min, pic_x_max),
  ybounds = c(pic_y_min, pic_y_max)
)

eyedens
```

Compute reinstatement: for each (participant, Background) compare the
recognition density map to the encoding density map for the same
participant + Background, and benchmark against a within-participant
permutation null. `eye_sim_diff` is the observed-minus-permuted similarity
in Fisher-z space — that's the per-pair effect.

```{r eyesim-reinstatement}
enc_dens <- eyedens |> filter(phase == "encoding")
rec_dens <- eyedens |> filter(phase == "recognition")

set.seed(1234)
reinstatement <- template_similarity(
  ref_tab      = enc_dens,
  source_tab   = rec_dens,
  match_on     = "pair_id",   
  permute_on = "participant", # match recognition→encoding for same pid+bg
  method       = "spearman",
  permutations = 1000
)

reinstatement
```

```{r eyesim-summary}
# Mean reinstatement (Fisher-z corrected against the within-participant null)
# overall and by Condition.
reinstatement |>
  summarise(
    mean_eye_sim      = mean(eye_sim,      na.rm = TRUE),
    mean_perm_sim     = mean(perm_sim,     na.rm = TRUE),
    mean_eye_sim_diff = mean(eye_sim_diff, na.rm = TRUE),
    n_pairs           = n()
  )

reinstatement |>
  group_by(Condition) |>
  summarise(
    mean_eye_sim_diff = mean(eye_sim_diff, na.rm = TRUE),
    sd_eye_sim_diff   = sd(eye_sim_diff,   na.rm = TRUE),
    n_pairs           = n(),
    .groups = "drop"
  )
```

# Write outputs

Persist the summary tables to `output/<group>/` as CSVs so they can be
shared, version-controlled, or fed into downstream stats scripts. All
filenames are prefixed with `group_label` (`TD`/`DS`) to keep groups
separate.

| file | what it is |
|---|---|
| `encoding_behavioral.csv` | per-trial encoding behavioral table |
| `recognition_behavioral.csv` | per-trial recognition behavioral table |
| `recognition_accuracy.csv` | HR / FAR / accuracy / d′ / c per participant |
| `recognition_accuracy_by_condition.csv` | same broken down by Condition |
| `recognition_duration_summary.csv` | back-task picture-display duration QC |
| `validation_encoding.csv`, `validation_recognition.csv` | calibration metrics per validation block |
| `missing_per_trial_{encoding,recognition}.csv` | per (participant, trial) NA-sample count + prop_missing in the 5-s window |
| `missing_per_subject_{encoding,recognition}.csv` | per-participant rollup of prop_missing across trials |
| `fixations_encoding.csv`, `fixations_recognition.csv` | I-VT fixation events |
| `fixation_aoi_summary_encoding.csv` | per (participant, trial, AOI) dwell |
| `fixation_summary_recognition.csv` | per-trial recognition fixation count + mean duration, by AOI (old + correct) |
| `fixation_summary_recognition_by_condition.csv` | participant × Condition × AOI rollup of the above |
| `emo_dwell_encoding.csv` | object-side dwell × emotion |
| `reinstatement.csv` | per-pair encoding↔recognition similarity |
| `reinstatement_by_condition.csv` | reinstatement summary × Condition |

```{r write-outputs}
out <- function(name) file.path(output_dir, paste0(group_label, "_", name))

# Behavioral
write_csv(encoding,    out("encoding_behavioral.csv"))
write_csv(recognition, out("recognition_behavioral.csv"))

# Recognition stats
write_csv(recognition_acc,              out("recognition_accuracy.csv"))
write_csv(recognition_acc_by_condition, out("recognition_accuracy_by_condition.csv"))
write_csv(recognition_duration_summary, out("recognition_duration_summary.csv"))

# Eye-tracking quality + events
write_csv(validation,             out("validation_encoding.csv"))
write_csv(recognition_validation, out("validation_recognition.csv"))

# Missing-data QC (per trial + per subject, both phases)
write_csv(trial_missing,                out("missing_per_trial_encoding.csv"))
write_csv(subject_missing,              out("missing_per_subject_encoding.csv"))
write_csv(recognition_trial_missing,    out("missing_per_trial_recognition.csv"))
write_csv(recognition_subject_missing,  out("missing_per_subject_recognition.csv"))

# Fixations
write_csv(fixations_behav,             out("fixations_encoding.csv"))
write_csv(recognition_fixations_behav, out("fixations_recognition.csv"))
write_csv(fix_summary,             out("fixation_aoi_summary_encoding.csv"))
write_csv(recognition_fix_summary,       out("fixation_summary_recognition.csv"))
write_csv(recognition_fix_by_condition,  out("fixation_summary_recognition_by_condition.csv"))
write_csv(emo_dwell,               out("emo_dwell_encoding.csv"))

# Gaze reinstatement
write_csv(reinstatement, out("reinstatement.csv"))

reinstatement_by_condition <- reinstatement |>
  group_by(Condition) |>
  summarise(
    mean_eye_sim      = mean(eye_sim,      na.rm = TRUE),
    mean_perm_sim     = mean(perm_sim,     na.rm = TRUE),
    mean_eye_sim_diff = mean(eye_sim_diff, na.rm = TRUE),
    sd_eye_sim_diff   = sd(eye_sim_diff,   na.rm = TRUE),
    n_pairs           = n(),
    .groups = "drop"
  )
write_csv(reinstatement_by_condition, out("reinstatement_by_condition.csv"))

list.files(output_dir)
```
